---
title: 'Known-biomass production models: the case study of the Bay the Biscay and
  Iberian Coast Ecoregion'
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: no
    toc: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Known-biomass production models (KBPMs) 

## Southern hake

```{r}
Data_hake<-read.csv("Hake_aux.csv",header = TRUE, sep= ";", encoding="UTF-8")

Data_Sh<-subset(Data_hake, Data_hake[,17] == "hke.27.8c9a")
min(Data_Sh$Year);max(Data_Sh$Year)


```


Surplus production and average SSB:

```{r}
B=Data_Sh$SSB; C=Data_Sh$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Sh=Data_Sh[-l,]
Data_Sh$SSB_aver=B_aver    
Data_Sh$SP=SP  
```

Fit model:

```{r}
model_Sh <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Sh)
summary(model_Sh)
```


Reference points:

```{r}

RF.Schaefer=function(model){
  
  a=coefficients(model)[1]
  b=coefficients(model)[2]
  
  B0<-as.numeric(-a/b)
  Bmsy<-as.numeric(B0/2)
  MSY<-as.numeric(a*Bmsy+b*Bmsy^2)
  x<-as.numeric(MSY/B0)
  Fmsy=as.numeric(a/2)
  
RP=c(round(B0),round(Bmsy),round(Fmsy,2),round(MSY),round(x,2))
RP=as.data.frame(RP)
rownames(RP)=c("SSB0","SSBmsy","Fmsy","MSY","MSY/SSB0")
return(RP)
}

RF.Schaefer(model_Sh)
```

Plot of the fit:



````{r}
a=coefficients(model_Sh)[1]
b=coefficients(model_Sh)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Sh$SSB_aver))
bv <- predict(model_Sh, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Sh$Year),max(Data_Sh$Year))
vec=min(df_aux$bv,Data_Sh$SP)
vec1=max(df_aux$bv,Data_Sh$SP)
last=dim(Data_Sh)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Southern hake",
                     ylim=c(vec,vec1))
points(Data_Sh$SSB_aver,Data_Sh$SP,pch=19,col="#56B4E9")
lines(Data_Sh$SSB_aver,Data_Sh$SP,lwd=2,col="#56B4E9")
points(Data_Sh$SSB_aver[1],Data_Sh$SP[1],pch=17,col="red",cex=1.5)
points(Data_Sh$SSB_aver[last],Data_Sh$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend("topleft", legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.75)



```

## Northern hake

```{r}
Data_hake<-read.csv("Hake.csv",header = TRUE, sep= ",", encoding="UTF-8")
Data_hake=Data_hake[,-1]
Data_Nh<-subset(Data_hake, Data_hake[,17] == "hke.27.3a46-8abd")
min(Data_Nh$Year);max(Data_Nh$Year)


```


Surplus production and average SSB:

```{r}
B=Data_Nh$SSB; C=Data_Nh$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Nh=Data_Nh[-l,]
Data_Nh$SSB_aver=B_aver    
Data_Nh$SP=SP  
```

Fit model:

```{r}
model_Nh <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Nh)
summary(model_Nh)
```


Reference points:

```{r}

RF.Schaefer(model_Nh)
```

Plot of the fit:



````{r}
a=coefficients(model_Nh)[1]
b=coefficients(model_Nh)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Nh$SSB_aver))
bv <- predict(model_Nh, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Nh$Year),max(Data_Nh$Year))
vec=min(df_aux$bv,Data_Nh$SP)
vec1=max(df_aux$bv,Data_Nh$SP)
last=dim(Data_Nh)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Northern hake",
                     ylim=c(vec,vec1))
points(Data_Nh$SSB_aver,Data_Nh$SP,pch=19,col="#56B4E9")
lines(Data_Nh$SSB_aver,Data_Nh$SP,lwd=2,col="#56B4E9")
points(Data_Nh$SSB_aver[1],Data_Nh$SP[1],pch=17,col="red",cex=1.5)
points(Data_Nh$SSB_aver[last],Data_Nh$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```


## Northern anglerfish

```{r}
Data_anglerfish<-read.csv("Anglerfish.csv",header = TRUE, sep= ",", encoding="UTF-8")
Data_anglerfish=Data_anglerfish[,-1]
Data_Na<-subset(Data_anglerfish, Data_anglerfish[,17] == "mon.27.78abd")
min(Data_Na$Year);max(Data_Na$Year)


```


Surplus production and average SSB:

```{r}
B=Data_Na$SSB; C=Data_Na$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Na=Data_Na[-l,]
Data_Na$SSB_aver=B_aver    
Data_Na$SP=SP  
```

Fit model:

```{r}
model_Na <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Na)
summary(model_Na)
```


Reference points:

```{r}

RF.Schaefer(model_Na)
```

Plot of the fit:



````{r}
a=coefficients(model_Na)[1]
b=coefficients(model_Na)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Na$SSB_aver))
bv <- predict(model_Na, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Na$Year),max(Data_Na$Year))
vec=min(df_aux$bv,Data_Na$SP)
vec1=max(df_aux$bv,Data_Na$SP)
last=dim(Data_Na)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Northern anglerfish",
                     ylim=c(vec,vec1))
points(Data_Na$SSB_aver,Data_Na$SP,pch=19,col="#56B4E9")
lines(Data_Na$SSB_aver,Data_Na$SP,lwd=2,col="#56B4E9")
points(Data_Na$SSB_aver[1],Data_Na$SP[1],pch=17,col="red",cex=1.5)
points(Data_Na$SSB_aver[last],Data_Na$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```



## Southern anglerfish

```{r}
Data_Sa<-subset(Data_anglerfish, Data_anglerfish[,17] == "mon.27.8c9a")
min(Data_Sa$Year);max(Data_Sa$Year)


```


Surplus production and average SSB:

```{r}
Data_Sa$catches=Data_Sa$landings
B=Data_Sa$SSB; C=Data_Sa$landings
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Sa=Data_Sa[-l,]
Data_Sa$SSB_aver=B_aver    
Data_Sa$SP=SP  
```

Fit model:

```{r}
model_Sa <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Sa)
summary(model_Sa)
```


Reference points:

```{r}

RF.Schaefer(model_Sa)
```

Plot of the fit:



````{r}
a=coefficients(model_Sa)[1]
b=coefficients(model_Sa)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Sa$SSB_aver))
bv <- predict(model_Sa, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Sa$Year),max(Data_Sa$Year))
vec=min(df_aux$bv,Data_Sa$SP)
vec1=max(df_aux$bv,Data_Sa$SP)
last=dim(Data_Sa)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Southern anglerfish",
                     ylim=c(vec,vec1))
points(Data_Sa$SSB_aver,Data_Sa$SP,pch=19,col="#56B4E9")
lines(Data_Sa$SSB_aver,Data_Sa$SP,lwd=2,col="#56B4E9")
points(Data_Sa$SSB_aver[1],Data_Sa$SP[1],pch=17,col="red",cex=1.5)
points(Data_Sa$SSB_aver[last],Data_Sa$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```



## Northern megrim

```{r}
Data_megrim<-read.csv("Megrim.csv",header = TRUE, sep= ",", encoding="UTF-8")
Data_megrim=Data_megrim[,-1]
Data_Nm<-subset(Data_megrim, Data_megrim[,17] == "meg.27.7b-k8abd")
min(Data_Nm$Year);max(Data_Nm$Year)


```


Surplus production and average SSB:

```{r}
Data_Nm$catches=Data_Nm$landings+Data_Nm$discards
B=Data_Nm$SSB; C=Data_Nm$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Nm=Data_Nm[-l,]
Data_Nm$SSB_aver=B_aver    
Data_Nm$SP=SP  
```

Fit model:

```{r}
model_Nm <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Nm)
summary(model_Nm)
```


Reference points:

```{r}

RF.Schaefer(model_Nm)
```

Plot of the fit:



````{r}
a=coefficients(model_Nm)[1]
b=coefficients(model_Nm)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Nm$SSB_aver))
bv <- predict(model_Nm, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Nm$Year),max(Data_Nm$Year))
vec=min(df_aux$bv,Data_Nm$SP)
vec1=max(df_aux$bv,Data_Nm$SP)
last=dim(Data_Nm)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Northern megrim",
                     ylim=c(vec,vec1))
points(Data_Nm$SSB_aver,Data_Nm$SP,pch=19,col="#56B4E9")
lines(Data_Nm$SSB_aver,Data_Nm$SP,lwd=2,col="#56B4E9")
points(Data_Nm$SSB_aver[1],Data_Nm$SP[1],pch=17,col="red",cex=1.5)
points(Data_Nm$SSB_aver[last],Data_Nm$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```



## Southern megrim

```{r}

Data_Sm<-subset(Data_megrim, Data_megrim[,17] == "meg.27.8c9a")
min(Data_Sm$Year);max(Data_Sm$Year)


```


Surplus production and average SSB:

```{r}
Data_Sm$catches=Data_Sm$landings+Data_Sm$discards
B=Data_Sm$SSB; C=Data_Sm$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_Sm=Data_Sm[-l,]
Data_Sm$SSB_aver=B_aver    
Data_Sm$SP=SP  
```

Fit model:

```{r}
model_Sm <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_Sm, algorithm = "port",
                lower=c(0,-Inf),upper=c(Inf,-0.0000226))
summary(model_Sm)
```


Reference points:

```{r}
RF.Schaefer(model_Sm)
```

Plot of the fit:



````{r}
a=coefficients(model_Sm)[1]
b=coefficients(model_Sm)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Sm$SSB_aver))
bv <- predict(model_Sm, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_Sm$Year),max(Data_Sm$Year))
vec=min(df_aux$bv,Data_Sm$SP)
vec1=max(df_aux$bv,Data_Sm$SP)
last=dim(Data_Sm)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Southern megrim",
                     ylim=c(vec,vec1))
points(Data_Sm$SSB_aver,Data_Sm$SP,pch=19,col="#56B4E9")
lines(Data_Sm$SSB_aver,Data_Sm$SP,lwd=2,col="#56B4E9")
points(Data_Sm$SSB_aver[1],Data_Sm$SP[1],pch=17,col="red",cex=1.5)
points(Data_Sm$SSB_aver[last],Data_Sm$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```



#  Identifying possible regime shifts



## Northern hake 

### 2017-2014-2011-2008-2005

```{r}
years=c(2017,2014,2011,2008,2005)

nR=5
lastyear=max(Data_Nh$Year)
modelretro=list()
vec=data.frame(matrix(0, ncol=5,nrow=nR))
names_retro=1:nR
for (i in 1:nR){
    newdf=subset(Data_Nh, Year<=(years[i]))
    fit <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=newdf)
    val=RF.Schaefer(fit)

    vec[i,]=t(val)
    print(summary(fit))
    modelretro[[i]]=fit
    names_retro[i]=paste(min(Data_Nh$Year), "-", years[i])

  }

  names(modelretro)=names_retro
  
  colnames(vec)=c("B0","B_MSY","F_MSY","MSY","MSYoverB0")
  rownames(vec)=names_retro

  vec

  # Plots Retrospective  ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

  bretro=list();vec_bv=0
  aretro=list()
  for (i in 1:nR){
    fit=modelretro[[i]]
    a=coefficients(fit)[1]
    b=coefficients(fit)[2];n=2
    cut=(-a/b)^(1/(n-1))
    av <- seq(0, cut, length.out = 3*length(Data_Nh$SSB_aver))
    bretro[[i]] <- predict(fit, newdata=(list(SSB_aver=av)))
    aretro[[i]]=av
    vec_bv=c(vec_bv,bretro[[i]])
}

a=coefficients(model_Nh)[1]
b=coefficients(model_Nh)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_Nh$SSB_aver))
bv <- predict(model_Nh, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)

  leg=c(min(Data_Nh$Year),max(Data_Nh$Year))
  vec=min(vec_bv,Data_Nh$SP,df_aux$bv)
  vec1=max(vec_bv,Data_Nh$SP,df_aux$bv)
  vec2=min(unlist(aretro),Data_Nh$SSB_aver,df_aux$av)
  vec3=max(unlist(aretro),Data_Nh$SSB_aver,df_aux$av)
  last=dim(Data_Nh)[1]

  plot(Data_Nh$SSB_aver,Data_Nh$SP,lwd=2,col="#56B4E9",xlab="SSB",ylab="SP",
                      main="Northern hake",
                 ylim=c(vec,vec1),type="l", xlim=c(vec2,vec3))
points(Data_Nh$SSB_aver,Data_Nh$SP,pch=19,col="#56B4E9")
points(Data_Nh$SSB_aver[1],Data_Nh$SP[1],pch=17,col="red",cex=1.5)
points(Data_Nh$SSB_aver[last],Data_Nh$SP[last],pch=15,col="red",cex=1.5)
lines(df_aux$av,df_aux$bv,col="black",lwd=2)

  for (i in 1:nR){
    lines(aretro[[i]],bretro[[i]],col=i+9,lwd=2)
  }

  colours=(1:nR)+9
  abline(h=0,lty=3,col="grey")
 legend("topleft",bg="transparent", legend=c(leg),
                   pch=c(17,15), cex=0.8,
                   box.lty=0,col="red",
                   x.intersp = 0.3, y.intersp = 0.75)
  legend("topright", bg="transparent",legend=c(paste(min(Data_Nh$Year), "-", lastyear),names_retro),
                   lwd=2, cex=0.8,
                   box.lty=0,col=c("black",colours),
                   x.intersp = 0.4, y.intersp = 0.75)




```


# Multi-species KBPMs


## South


```{r}

Years_min=c(min(Data_Sa$Year), min(Data_Sh$Year), min(Data_Sm$Year))
Years_max=c(max(Data_Sa$Year), max(Data_Sh$Year), max(Data_Sm$Year))
min=max(Years_min);min
max=min(Years_max);max

Data_Sa_sub=subset(Data_Sa,Data_Sa$Year<=max & Data_Sa$Year>=min)
Data_Sh_sub=subset(Data_Sh,Data_Sh$Year<=max & Data_Sh$Year>=min)
Data_Sm_sub=subset(Data_Sm,Data_Sm$Year<=max & Data_Sm$Year>=min)

SSB_add=rowSums(data.frame(Data_Sa_sub$SSB, Data_Sh_sub$SSB, Data_Sm_sub$SSB))
Catches_add=rowSums(data.frame(Data_Sa_sub$catches, Data_Sh_sub$catches, Data_Sm_sub$catches))
Data_south=data.frame(SSB=SSB_add, catches=Catches_add, Year=min:max)

min(Data_south$Year);max(Data_south$Year)


```


Surplus production and average SSB:

```{r}
B=Data_south$SSB; C=Data_south$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_south=Data_south[-l,]
Data_south$SSB_aver=B_aver    
Data_south$SP=SP  
```

Fit model:

```{r}
model_south <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_south)
summary(model_south)
```


Reference points:

```{r}
RF.Schaefer(model_south)
```

Plot of the fit:



````{r}
a=coefficients(model_south)[1]
b=coefficients(model_south)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_south$SSB_aver))
bv <- predict(model_south, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_south$Year),max(Data_south$Year))
vec=min(df_aux$bv,Data_south$SP)
vec1=max(df_aux$bv,Data_south$SP)
last=dim(Data_south)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="South",
                     ylim=c(vec,vec1))
points(Data_south$SSB_aver,Data_south$SP,pch=19,col="#56B4E9")
lines(Data_south$SSB_aver,Data_south$SP,lwd=2,col="#56B4E9")
points(Data_south$SSB_aver[1],Data_south$SP[1],pch=17,col="red",cex=1.5)
points(Data_south$SSB_aver[last],Data_south$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```


## North


```{r}

Years_min=c(min(Data_Na$Year), min(Data_Nh$Year), min(Data_Nm$Year))
Years_max=c(max(Data_Na$Year), max(Data_Nh$Year), max(Data_Nm$Year))
min=max(Years_min);min
max=min(Years_max);max

Data_Na_sub=subset(Data_Na,Data_Na$Year<=max & Data_Na$Year>=min)
Data_Nh_sub=subset(Data_Nh,Data_Nh$Year<=max & Data_Nh$Year>=min)
Data_Nm_sub=subset(Data_Nm,Data_Nm$Year<=max & Data_Nm$Year>=min)

SSB_add=rowSums(data.frame(Data_Na_sub$SSB, Data_Nh_sub$SSB, Data_Nm_sub$SSB))
Catches_add=rowSums(data.frame(Data_Na_sub$catches, Data_Nh_sub$catches, Data_Nm_sub$catches))
Data_north=data.frame(SSB=SSB_add, catches=Catches_add, Year=min:max)

min(Data_north$Year);max(Data_north$Year)


```


Surplus production and average SSB:

```{r}
B=Data_north$SSB; C=Data_north$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_north=Data_north[-l,]
Data_north$SSB_aver=B_aver    
Data_north$SP=SP  
```

Fit model:

```{r}
model_north <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_north)
summary(model_north)
```


Reference points:

```{r}


RF.Schaefer(model_north)
```

Plot of the fit:



````{r}
a=coefficients(model_north)[1]
b=coefficients(model_north)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_north$SSB_aver))
bv <- predict(model_north, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_north$Year),max(Data_north$Year))
vec=min(df_aux$bv,Data_north$SP)
vec1=max(df_aux$bv,Data_north$SP)
last=dim(Data_north)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="North",
                     ylim=c(vec,vec1))
points(Data_north$SSB_aver,Data_north$SP,pch=19,col="#56B4E9")
lines(Data_north$SSB_aver,Data_north$SP,lwd=2,col="#56B4E9")
points(Data_north$SSB_aver[1],Data_north$SP[1],pch=17,col="red",cex=1.5)
points(Data_north$SSB_aver[last],Data_north$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```



## Global


```{r}
Years_min=c(min(Data_north$Year), min(Data_south$Year))
Years_max=c(max(Data_north$Year), max(Data_south$Year))
min=max(Years_min);min
max=min(Years_max);max

Data_south_sub=subset(Data_south,Data_south$Year<=max & Data_south$Year>=min)
Data_north_sub=subset(Data_north,Data_north$Year<=max & Data_north$Year>=min)

SSB_add=rowSums(data.frame(Data_south_sub$SSB,Data_north_sub$SSB))
Catches_add=rowSums(data.frame(Data_south_sub$catches,Data_north_sub$catches))
Data_global=data.frame(SSB=SSB_add, catches=Catches_add, Year=min:max)

min(Data_global$Year);max(Data_global$Year)


```


Surplus production and average SSB:

```{r}
B=Data_global$SSB; C=Data_global$catches
SP=B[-1]; l=length(B); B_aver=B[-1]
    for (i in 1:(l-1)){
      SP[i]=B[i+1]-B[i]+C[i]
      B_aver[i]=(B[i+1]+B[i])/2
    }
    B=B_aver
# Delete last year
Data_global=Data_global[-l,]
Data_global$SSB_aver=B_aver    
Data_global$SP=SP  
```

Fit model:

```{r}
model_global <- nls(SP~a*SSB_aver + b*SSB_aver^2, start=list(a=1.4, b=-0.0000226), data=Data_global)
summary(model_global)
```


Reference points:

```{r}

RF.Schaefer(model_global)
```

Plot of the fit:



````{r}
a=coefficients(model_global)[1]
b=coefficients(model_global)[2];n=2
cut=(-a/b)^(1/(n-1))
av <- seq(0, cut, length.out = 3*length(Data_global$SSB_aver))
bv <- predict(model_global, newdata=(list(SSB_aver=av)))

df_aux=data.frame(av,bv)


leg=c(min(Data_global$Year),max(Data_global$Year))
vec=min(df_aux$bv,Data_global$SP)
vec1=max(df_aux$bv,Data_global$SP)
last=dim(Data_global)[1]
plot(df_aux$av,df_aux$bv, type="l",
                     col="black",lwd=2,xlab="SSB",ylab="SP",
                     main="Global",
                     ylim=c(vec,vec1))
points(Data_global$SSB_aver,Data_global$SP,pch=19,col="#56B4E9")
lines(Data_global$SSB_aver,Data_global$SP,lwd=2,col="#56B4E9")
points(Data_global$SSB_aver[1],Data_global$SP[1],pch=17,col="red",cex=1.5)
points(Data_global$SSB_aver[last],Data_global$SP[last],pch=15,col="red",cex=1.5)
abline(h=0,lty=3,col="grey")
legend(c(vec,vec1), legend=leg,bg="transparent",
                       pch=c(17,15), cex=0.7,
                       box.lty=0,col="red",x.intersp = 0.3, y.intersp = 0.65)



```


#  Testing the environmental variability om stocks’ status

## Selecting lag

```{r}
years=Data_Nh$Year
SP=Data_Nh$SP
nlag=4
# NAO ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~


NAO<-read.csv("NAO.csv",header = TRUE, sep = ";", encoding="UTF-8")
y_env=NAO$Ano

  f_year=years[1]
  l_year=years[length(years)]
  data_env=data.frame(SP,years)
  ind=which(y_env==f_year)
  ind1=which(y_env==l_year)
  data_env$NAO=NAO$Mean[ind:ind1]
  a=as.data.frame(matrix(0,nrow=nrow(data_env),ncol=nlag))
  data_env=cbind(data_env,a)
 vec_env=1:nlag
  for (i in 1:nlag){
    ind=which(y_env==(f_year-i))
    ind1=which(y_env==(l_year-i))
 
  data_env[,i+3]=NAO$Mean[ind:ind1]
  vec_env[i]=paste("NAO lag",i)
  colnames(data_env)=c("SP","years","NAO lag 0",vec_env)
  }
  aux=c("SP","years","NAO lag 0",vec_env)

# AMO ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
AMO<-read.csv("AMO.csv",header = TRUE, sep = ";", encoding="UTF-8")
AMO$Mean<-apply(AMO[,-1],1, mean)
y_env=AMO$Ano

  f_year=years[1]
  l_year=years[length(years)]
  ind=which(y_env==f_year)
  ind1=which(y_env==l_year)
  data_env$AMO=AMO$Mean[ind:ind1]
a=as.data.frame(matrix(0,nrow=nrow(data_env),ncol=nlag))
data_env=cbind(data_env,a)
vec_env=1:nlag
  for (i in 1:nlag){
    ind=which(y_env==(f_year-i))
    ind1=which(y_env==(l_year-i))
  
  data_env[,i+8]=AMO$Mean[ind:ind1]
    vec_env[i]=paste("AMO lag",i)
  colnames(data_env)=c(aux,"AMO lag 0",vec_env)
}

library(corrplot)
cor.mtest <- function(mat, ...) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat<- matrix(NA, n, n)
  diag(p.mat) <- 0
  for (i in 1:(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], ...)
      p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
    }
  }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
p.mat <- cor.mtest(data_env[,-2])
corrplot(round(cor(as.matrix(data_env[,-2]),use="na.or.complete"),2),method="number",diag=T,p.mat = p.mat, sig.level = 0.05, type = "lower", number.cex = 0.6, tl.cex=0.4, tl.pos="d")


```


## Fitting

```{r}
Data_Nh$i=data_env$`AMO lag 4`
```

### Multiplicative



```{r}

model_M <- nls(SP~exp(1)^{c*i}*(a*SSB_aver + b*SSB_aver^2), start=list(a=1.4, b=-0.0000226,c=1), data=Data_Nh)
summary(model_M)

```

### Additive

```{r}

model_A<- nls(SP~a*SSB_aver + b*SSB_aver^2+c*i, start=list(a=1.4, b=-0.0000226,c=1), data=Data_Nh)

summary(model_A)
```


### Plot and error measures


```{r}
bv <- predict(model_Nh)
bv1 <- predict(model_M)
bv2 <- predict(model_A)


MAPE=function(predicted, real){ (1/length(real))*(sum(abs((predicted-real)/real)))  }
RMSE=function(predicted, real){sqrt(sum((predicted-real)^2)/length(real))}

RMSE(bv1,Data_Nh$SP);RMSE(bv2,Data_Nh$SP);RMSE(bv,Data_Nh$SP)
MAPE(bv1,Data_Nh$SP);MAPE(bv2,Data_Nh$SP);MAPE(bv,Data_Nh$SP)
AIC(model_M);AIC(model_A);AIC(model_Nh)

leg=c("Observed","Base model","Environmental Multiplicative","Environmental Additive")

vec=min(c(bv,bv1,bv2,Data_Nh$SP))
vec1=max(c(bv,bv1,bv2,Data_Nh$SP))
plot(Data_Nh$Year,Data_Nh$SP,lwd=2,col="#56B4E9",xlab="Years",ylab="SP",
                 main="Environmental fits",
                 ylim=c(vec,vec1),type="b",pch=19)
lines(Data_Nh$Year,bv,col=1,lwd=2)
lines(Data_Nh$Year,bv1,col=2,lwd=2)
lines(Data_Nh$Year,bv2,col=3,lwd=2)
abline(h=0,lty=3,col="grey")

legend("topleft",bg="transparent", legend=c(leg),
                   lwd=c(NA,2,2,2), pch=c(19,NA,NA,NA), cex=0.7,
                   box.lty=0,col=c("#56B4E9",1:3))

```